---
title: "Raptors project"
output: pdf_document
date: "2026-02-22"
---

```{r}
library(tidyverse)
library(broom)
library(janitor)
library(knitr)
```

```{r}
nba = read_csv("games-2.csv")

#filter the raptors 
raptors = nba %>%
  filter(HOME_TEAM_ID == 1610612761 | VISITOR_TEAM_ID == 1610612761)

```

```{r}
#cleaning the data
raptors = raptors %>%
  mutate(
    team_points = ifelse(HOME_TEAM_ID == 1610612761, PTS_home, PTS_away),
    opp_points  = ifelse(HOME_TEAM_ID == 1610612761, PTS_away, PTS_home),
    home = ifelse(HOME_TEAM_ID == 1610612761, 1, 0),
    win = ifelse(team_points > opp_points, 1, 0)
  )

raptors <- raptors %>%
  filter(!is.na(team_points),
         !is.na(opp_points))
```

```{r}
raptors <- raptors %>%
  mutate(win = ifelse(team_points > opp_points, 1, 0))
mean(raptors$win, na.rm = TRUE)
```
The Raptors won approximately 51.5% of games across the observed seasons. This means they were slightly above average overall. It gives a baseline understanding of how competitive the team was during the seasons analyzed.

```{r}
#home court advantage
raptors %>%
  group_by(home) %>%
  summarise(win_rate = mean(win))
```
The Raptors won more games at home than away.This shows that playing at home gives the team an advantage. Possible reasons include:support from fans, less travel fatigue, familiarity with the arena etc. 

```{r}
ggplot(raptors, aes(x = team_points, fill = factor(win))) +
  geom_density(alpha = 0.5)
```
The density plot showed that when the Raptors scored more points, they were much more likely to win.
This confirms something expected but important: scoring more leads to winning.

```{r}
#logistics regression model
raptors <- raptors %>%
  mutate(
    team_fg_pct = ifelse(HOME_TEAM_ID == 1610612761,
                         FG_PCT_home,
                         FG_PCT_away),

    team_fg3_pct = ifelse(HOME_TEAM_ID == 1610612761,
                          FG3_PCT_home,
                          FG3_PCT_away),

    team_reb = ifelse(HOME_TEAM_ID == 1610612761,
                      REB_home,
                      REB_away)
  )

model = glm(win ~ + team_fg_pct + team_reb + home,
             data = raptors,
             family = "binomial")

summary(model)

```
FGP(field goal point) was the strongest factor.
When the Raptors shoot more accurately, their chances of winning increase significantly.
In simple words:
Better shooting = much higher chance of winning.
Even small improvements in shooting percentage make a noticeable difference. When the Raptors grab more rebounds, their probability of winning increases.Playing at home increases the probability of winning, even after accounting for shooting and rebounds.
This confirms that home court matters.The Raptors have a real advantage when they play at home.
```{r}
#predict 
raptors$predicted_win_prob <- predict(model, type = "response")
ggplot(raptors, aes(x = team_points, y = predicted_win_prob)) +
  geom_point(alpha = 0.4) +
  geom_smooth()
```
The plot shows that as the Raptors score more points, the predicted probability of winning increases.
In simple terms:
More points → higher chance of winning.
The upward curve confirms that scoring performance strongly influences win probability.
